Arabic Entity Graph Extraction using Morphology, Finite State Machines, and Graph Transformations

### 2018-04-12 tex/cicling.tex
changes to answer, colors, and some reviewers addressed

%\title{Instructions for NAACL HLT 2010 Proceedings\Thanks{This...}}
\title{Detection of Arabic Entity Graphs using Morphology, Finite State Machines, and Graph Transformations}

%\author{ Jad Makhlouta \\\And
%Hamza Harkous \\
%  American University of Beirut \\
%  {\tt \{jem04, hhh20, fz11\}@aub.edu.lb } \\\And 
%  Fadi Zaraket 
%}
\date{}

@@

\begin{abstract}
Research on automatic recognition of named entities
%events and relations amongst them 
from Arabic text uses techniques that work well for the 
Latin based languages such as local grammars, 
statistical learning models, pattern matching, 
and rule-based techniques. 
These techniques boost their results by using application 
specific 
corpora, parallel language corpora,
and morphological stemming analysis. 
We propose a method for extracting entities, events, 
and relations amongst them
from Arabic text using 
a hierarchy of finite state machines driven by morphological 
features such as part of speech and gloss tags,
and graph transformation algorithms. 
We evaluated our method on three natural language processing 
applications.
We automated the extraction of narrators and 
narrator relations from 
several corpora of Islamic narration books (hadith). 
We automated the extraction of genealogical family trees 
from Biblical texts. 
Finally, we automated locating and extracting
individual biographies from historical biography books.
In all applications, our method reports high precision and recall
and learns lemmas about phrases that improve results. 
\end{abstract}

@@

Industrial tools that perform NE extraction for Arabic exist~\cite{Sak09,IdentiFinder,basis:06,ANEE:07},
but little is said about the techniques behind them and no formal evaluation of the tools exist. 
Researchers proposed and evaluated 
local grammars with morphological stemming~\cite{ZAGHOUANI10,Traboulsi:09},
rule-based systems powered with local grammars~\cite{ShaalanR09} 
and stop words ~\cite{Abuleil04},
statistical analysis boosted with Arabic-English parallel corpora~\cite{BenajibaZDR10}, 
statistical analysis boosted with task specific gazetteers~\cite{ANERSys:07,BenajibaDR08},
and pattern matching powered with morphological stemming~\cite{jumaily2011,TAGARAB98}
to perform Arabic NER.
We discuss these techniques and compare to them in 
the ``Related work'' Section.

In this paper, we consider the problem of NER
in the context of more complex natural language
processing (NLP) tasks that look for NEs and 
relations amongst the detected NEs and build entity graphs.
An {\em entity graph} is a graph where NEs are nodes and relational entities (RE) are edges.
For example, consider the tasks of extracting a time table
from a document describing the events of a conference, 
extracting genealogical family trees based on person names 
from biblical text,  and
extracting routes and maps from travel itineraries.
Also consider applications with cross-document NER
such as extracting narrator authentication graphs based on narration 
documents that include narrator sequences, and biography documents 
that include credibility information about the narrators. 
We hypothesize that NEs of the same kind happen 
at proximity of each other, and we base our method on that 
working hypothesis.
We target NLP applications and corpora where our working hypothesis
obviously holds.
These applications help build databases of NEs
that we can later use in other related contexts.

We propose a method for extracting NEs and 
relations amongst entities using 
morphological features such as part of speech (POS) and gloss tags, 
a hierarchy of finite state machines, 
and graph transformations such as node merge, node split, and replace edge transformations. 
Morphological analysis helps reduce ambiguities and resolves lexical items with their morphological variations.
For each application, we augment the lexicon of an inhouse morphological 
analyzer with lexical elements and tags that describe categories relevant to the application. 
We also build finite state transducers (FST) that take 
morphological tags as input and detect the NEs.
Manually built FSTs can express rules that are not intuitive (and sometimes impossible) 
to express with local grammars using regular expressions.
Consider, for example, counting the number of tags up to a threshold, or capturing left and right recursive structures 
which are common in Arabic since an Arabic phrase-sentence can start with a name or a verb.
We finally infer relations between the NEs based on their context.

@@

We evaluated our method and used it to solve three 
NLP applications including extracting genealogical family
trees from Biblical text, extracting a graph of 
narrators from hadith books, and extracting biographies from biography books.
Our results show high precision and recall metrics and report
learning interesting named entities and relational lemmas.
In this paper we make the following contributions.
\begin{itemize}
\item We present a method for named entity and entity graph
extraction from Arabic text using morphology, state 
machines, and graph transformation algorithms.
\item We present an Arabic lexicon augmented 
with (1) person names collected from several resources as well as
learned from our corpora, and (2) tags suitable to extract complex names,
kinship, and narration sequence relations. 
\item We learn semantics for morphological features associated 
with detected phrases that are equivalent to predefined features
modulo one or two graph transformations. 
The learned semantics improved recall and precision by 
5 and 6\%, respectively in the genealogy application.
\item We perform Arabic cross-document named entity and 
relational entity extraction using graph algorithms. 
The use of a graph generated from the hadith application 
improved precision in the biography application from 40\% to 93\%.
%\item We present a novel equality metric to compare Arabic 
%names based on name structure and morphological features. 
\end{itemize}

@@

Our method takes a set of Arabic text documents and builds 
a graph where named entities are nodes and edges are relations 
between the nodes. 
Our target graph has a predefined set of relations expressed 
as edge labels such as ``parent'' and ``spouse'' where the nodes
are person names.
We use an inhouse morphological analyzer augmented with 
a list of lexicon elements relevant to the application.
We also augment the analyzer with
a set of tags that classify the 
elements into several categories 
relevant to the targeted named entities and edge labels.

The method passes the input text string to the morphological 
analyzer. The analyzer processes the input text and whenever
it identifies a morpheme, it calls the application morpheme
classifier with the morpheme, the tags of the morpheme, 
and the current solution context.
Note that several morphological solutions may exist for the same
input string.
The morpheme classifier either stores the morpheme in a list of
unresolved morphemes, 
or resolves the unresolved morphemes and produces
a category as input to the application FST.
The application FST is manually built to
detect {\bf (1)} named entities,  
{\bf (2)} sequences of named entities,  
{\bf (3)} predefined edges connecting named entities, 
{\bf (4)} previously unknown named entities occurring between predefined edges, 
and {\bf (5)} previously unknown edge labels occurring between named entities.
%Note that there might be more than one level of finite state
%transducers to handle an NLP application.
The detected named entities and edges form an entity graph.

Finally, the method uses a morphological and structural distance
function to compute equality between the named entities and
to consequently merge the nodes of the graph that represent 
equal named entities.
%When nodes are merged, the edge labels are augmented to be the 
%concatenation of old edges.
The method then learns semantics for the discovered edge labels 
and suggests edge replacement algorithms to transform the 
discovered edge labels into the predefined edges. 
The method also interactively learns the semantics of other 
detected edge labels that it could not infer 
in terms of the predefined set of relations.

@@

Due to religious and political reasons, writing the traditions was forbidden 
until the days of the Umayad Calif,
\novocalize
\RL{`mr bn `bd al`zyz}(717-720 AC), seventy or so years after the prophet. 
Many inconsistencies were introduced to the literature which necessitates 
a thorough authentication study before the use of a tradition in jurisprudence.
\vocalize
%An Islamic hadith scholar studies the authenticity of 
%the \noArRL{sanad} of a 
%set of related traditions before using these traditions 
%jurisprudence. 
%While different Islamic schools of thought differ on 
%how to interpret the content, they almost all agree
All Islamic schools of thought agree
that scholars can not use a tradition whose sanad lacks authenticity for jurisprudence.

The authenticity of a \noArRL{.hady_t} depends on 
the credibility of the narrators as reported in 
separate biography books. 
The study of \noArRL{.hady_t} authentication is 
currently manual and error prone due to the huge number
of existing traditions and tradition books. 
%Hadithopedia~\cite{Hadithopaedia:08}
%reports that a non-comprehensive collection of tradition 
%and biography books amounts to more than 300 thousand lines of 
%text. 
Al-Azami\cite{Al-Azami-91} cites more than eleven books
of digitized tradition books each of several volumes, and a dozen
other biography and secondary authentication books such
as a geographical dictionary of places in hadith.

@@

The hadith FST generates sequences of hadith narrators
such as $\langle n_1,n_2,n_3,n_4,n_5 \rangle$ shown in 
Figure~\ref{f:exhadith}. 
Our target is to automate the extraction of 
narrator graphs from one or more narration books
similar to the directed acyclic graph (DAG) 
in Figure~\ref{f:hadith}(b).
%We automatically extracted the POR from the three books of 
%hadith~\cite{IbnHanbal,AlKulayni,AlTousi}.
The nodes in boxes are the \RL{matn} of the hadith, 
and the other nodes are the narrators.

@@

\subsection{Narrator distance metric}
%MultiKey Hashing
%explain what are the keys used for each value and why suitable
% and say that we use the computed keys also conduct the 
% biography lookup
Two narrator names may refer to the same person but may 
differ in text and in structure. 
For example, the names \noTrutfRL{عبد الله بن عمر}, 
\noTrutfRL{ابن عمر}, and \noTrutfRL{ابن عمر رضي الله عنه}
refer to the same person. 
Our metric uses the \noTrutfRL{ابن} (son) name connector and 
its morphological equivalents to order the person names
that appear within a narrator entity. 
The metric compares the name hierarchy and if the hierarchy 
does not match, as in the case of 
\noTrutfRL{عبيد الله بن عبد الله بن عمر}  and his parent 
\noTrutfRL{عبد الله بن عمر},
the metric returns \cci{false}.
The metric also compares name qualifiers and if two name 
qualifiers of the same morphological POS tag,  
such as the place possessive qualifiers 
\utfRL{العراقي} (the Iraqi) and \utfRL{المصري} (the Egyptian) 
conflict, the metric returns \cci{false}.
The metric then compares the morphological stems of the person 
names that are in order, ignores the ones that are not in order,
and computes the fraction of the matching names.

Both our metric and the Levenshtein distance metric 
used in~\cite{Azmi-2010}
achieved 100\% precision on an arbitrarily selected set that 
included 208 narrator pairs with 102 equal narrators.
Our metric achieved 95\% recall versus 72\% recall for the 
Levenshtein distance metric. 
Which means that both metrics do not announce names equal if they
are not, but our metric is better at detecting equal names.
We believe that our metric is adequate to compare 
complex Arabic names including contemporary ones in general.

\section{The biography application}
Biography books include one or more short biography 
for each hadith narrator. 
The short biography may contain information 
about the dates and locations of birth and death, 
the professors and the students of the narrator,
and one or more evaluations of the credibility of the narrator.

We use a similar approach to the hadith application
to locate narrator names within the biography books. 
A biography text may contain the name of the narrator described
in the biography as well as the names of other narrators who happen
to be the professors, students, or the persons who evaluated the 
narrator.
Our method uses the narrator graph 
extracted from several hadith books 
as described earlier to segment the biographies
and detect the owner of each biography.

Once our method detects a narrator, it colors the node in  the 
narrator graph corresponding to it. 
The professors and students of the narrator
are the neighbors of the 
node corresponding to the narrator in the graph. 
Consequently, whenever a detected narrator entity 
moves away from the 
colored local cluster in the narrator graph, 
we change the color and infer that we moved to a separate
biography.
The owner of each biography is the narrator at the center of the
colored graph.

The naive implementation of this approach is computationally
expensive as it requires searching for each discovered narrator
in the narrator graph. 
We implemented a multi-hash narrator technique to 
improve narrator localization in the graph and in the biography 
books where we associate
each narrator with all narrator names that can be generated 
from its structure and compute a hash key and a relevance score 
for that association. 
We omit the discussion for brevity and we will report on that 
elsewhere.

Scholars can use the located biographies to annotate
each narrator with corresponding biographies thus automating
the task of hadith authentication checks. 
They can also annotate the narrator graph narrator graph with the locations and 
the lifetime of the narrators and 
check for time and location overlap that are necessary conditions
for the consistency of two 
neighboring narrators. 
Scholars can also use the method to check the correspondence
of different biography and hadith books.

We believe that this is the first work that applies
{\em Arabic cross-document NLP}, and we believe
that this approach can be extended to several applications
such as the cross analysis of (1) investigation versus 
interrogation reports, and (2) insurance claims versus 
accident police reports to find inconsistencies.

@@

Our method takes as input Arabic text with a genealogical
list such as the genealogical list in
Figure~\ref{f:genelists}(a) 
and automatically extracts the genealogical family tree 
where person names are entities and edges represent 
``spouse'' and ``child'' labels such as 
Figure~\ref{f:genelists}(b).
We marked the person names with bars above them, and we 
filled the spouse nodes with grey. 
To perform this task, our method first passes the text to
the morphological analyzer that reports detected morphemes 
with their corresponding tags.

We augmented the tag set of our morphological analyzer with 
the following categories relevant to the genealogical 
application. 
\begin{itemize}
\item {\em NAME} a person name that can be either male or female. 509 lexicon items. 
\item {\em CHILD} a connector with a descent semantics such as \utfRL{ابن} (son). 12 lexicon items. 
\item {\em CHILDREN} a connector with a plural descent semantics such as \utfRL{ابناء} (children). 16 lexicon items. 
\item {\em PARENT} a connector with parenthood semantics such as \utfRL{أباه} (his father). 7 lexicon items. 
\item {\em SPOUSE} a connector with marital partnership semantics such as \utfRL{سريّته} (his concubine). 11 lexicon items. 
\item {\em SPOUSES} a connector with plural partnership semantics such as \utfRL{سراري} (his concubines). 7 lexicon items. 
\item {\em SIBLING} a connector with sisterhood semantics such as \utfRL{أخيه} (his brother). 14 lexicon items.
\item {\em SIBLINGS} a connector with plural sisterhood semantics such as \utfRL{اخوته} (his siblings). 18 lexicon items. 
\end{itemize}
Note that (1) some lexicon items might be categorized into 
more than one of the above categories, and 
(2) the method infers the gender from POS and gloss tags 
of the person name as well as from those of the 
other categories. 
In case of conflict, the method gives precedence to the 
gender inferred from the person names. 
Note also that the phrases
\utfRL{اسحاق بن يعقوب} (Izhac - son of - Jacob)
and 
\utfRL{اسحاق ابنه يعقوب} (Izhac - his son - Jacob)
contain the same named entities, and are identical 
as far as stemming is concerned. 
However, the analysis 
of the suffix morphemes of \utfRL{ابنه} results in inferring
a parent category in the second phrase.

@@

ANERSys boosts its capabilities to detect named entities by using 
task specific corpora and gazetteers and ignores 
POS tags~\cite{ANERSys:07,BenajibaDR08}. 
We are similar to ANERSys in that we augment the lexicon of our
morphological analyzer with application relevant lexicon items
and tags. 
We differ in that they use statistical analysis models such as
n-gram and maximum entropy measures.
In other work, they
achieve major improvements when they use lexical and
morphosyntactic features and a parallel Arabic and English 
corpora to bootstrap noisy features~\cite{BenajibaZDR10}.

@@

We evaluated our method against three books of 
hadith~\cite{IbnHanbal,AlTousi,AlKulayni},
the Genesis~\cite{},
and one book of biographies with several volumes~\cite{khoei}. 
We define our metrics, present our evaluation reference with interannotation agreement results, 
and then present the results for the hadith
biography, and genealogical list applications.

@@

{\bf Biography narrator detection.}~~
Table~\ref{t:narratorDifference} shows the accuracy results for the narrator detection and the biography boundary detection
for the Khoei biography book.
The first row shows the results computed by considering the biography book alone without the help of the narrator graph
extracted from the hadith books. 
Without the narrator graph, our method detected 94\% of the narrators mentioned in the book with 81\% precision. 
The method without the narrator graph, detected correctly (94\% precision) the boundaries of the biographies.

With the help of the narrator graph, our method detected 65\% of the narrator names in the biography book. 
The rest of the names were simply not found in the hadith books that we used to generate the narrator graph. 
With the use of the narrator graph and the coloring algorithm, our method detected the boundaries of the biographies
with 93\% precision. 
We believe that an exhaustive graph that includes narrator names from more books of hadith will improve the 65\% narrator 
detection recall metric.
In all cases, narrator names that exist in the biography and not in the hadith books are not of interest to the narrator
graph annotation problem.

%The numbers in the right hand side of Table~\ref{t:narratorDifference}  may explain why 
Narrator detection in biography books is harder than in hadith books. 
For example, the ratio of narrator names to the rest of the text is 22\% in Musnad while it is 49\% in Khoei,
and the average length of a narrator sequence in Musnad is 2.6 narrators while it is 3.5 narrators in Khoei.
With the use of the narrator graph extracted from the hadith books, our method was able to improve biography boundary detection
from 41\% to 93\%.
This is evidence that cross-document entity detection helps to accurately find named entities in harder NLP tasks.

\begin{table*}[bt]                                     
\centering                                            
\caption{Accuracy results for the biography application.}
\resizebox{.7\columnwidth}{!}{

%\begin{tabular}{cp{0.3cm}c}

>>>>

%\title{Instructions for NAACL HLT 2010 Proceedings\Thanks{This...}}
\title{Arabic Entity Graph Extraction using Morphology, Finite State Machines, and Graph Transformations}

\author{ Jad Makhlouta\inst{1} \and
    Fadi Zaraket\inst{1} \and 
Hamza Harkous\inst{1} }
\institute{ American University of Beirut \\
  \email{ \{jem04, fz11, hhh20\}@aub.edu.lb } }
\date{}

@@

\begin{abstract}
Research on automatic recognition of named entities
%events and relations amongst them 
from Arabic text uses techniques that work well for the 
Latin based languages such as local grammars, 
statistical learning models, pattern matching, 
and rule-based techniques. 
These techniques boost their results by using application 
specific 
corpora, parallel language corpora,
and morphological stemming analysis. 
We propose a method for extracting entities, events, 
and relations amongst them
from Arabic text using 
a hierarchy of finite state machines driven by morphological 
features such as part of speech and gloss tags,
and graph transformation algorithms. 
We evaluated our method on two natural language processing 
applications.
We automated the extraction of narrators and 
narrator relations from 
several corpora of Islamic narration books. 
We automated the extraction of genealogical family trees 
from Biblical texts. 
%Finally, we automated locating and extracting
%individual biographies from historical biography books.
In all applications, our method reports high precision and recall
and learns lemmas about phrases that improve results. 
\end{abstract}

@@

Industrial tools that perform NE extraction for Arabic exist~\cite{Sak09,IdentiFinder,basis:06,ANEE:07},
but little is said about the techniques behind them and no formal evaluation of the tools exist. 
Researchers proposed and evaluated 
local grammars with morphological stemming~\cite{ZAGHOUANI10,Traboulsi:09},
rule-based systems powered with local grammars~\cite{ShaalanR09} 
and stop words ~\cite{Abuleil04},
statistical analysis boosted with Arabic-English parallel corpora~\cite{BenajibaZDR10}, 
statistical analysis boosted with task specific gazetteers~\cite{Benajiba:07,Benajiba:08},
and pattern matching powered with morphological stemming~\cite{jumaily2011,TAGARAB98}
to perform Arabic NER.
We discuss these techniques and compare to them in 
the ``Related work'' Section.

In this paper, we consider the problem of NER
in the context of more complex natural language
processing (NLP) tasks that look for NEs and 
relations amongst the detected NEs and build entity graphs.
An {\em entity graph} is a graph where NEs are nodes and relational entities (RE) are edges.
For example, consider the tasks of 
extracting genealogical family trees based on person names 
from biblical text, 
extracting routes and maps from travel itineraries,
and 
extracting narrator graphs based on narration 
documents that include narrator sequences.
%, and biography documents 
%that include credibility information about the narrators. 
We hypothesize that NEs of the same kind happen 
at proximity of each other, and we base our method on that 
working hypothesis.
We target NLP applications and corpora where our working hypothesis
obviously holds.
These applications help build databases of NEs
that we can later use in other related contexts.

We propose a method for extracting NEs and 
relations amongst entities using 
morphological features such as part of speech (POS) and gloss tags, 
a hierarchy of finite state machines, 
and graph transformations such as node merge, node split, and replace edge transformations. 
Morphological analysis helps reduce ambiguities and resolves lexical items with their morphological variations.
For each application, we augment the lexicon of Sarf~\cite{ATMine09}, an inhouse morphological analyzer, 
with lexical elements and tags that describe categories relevant to the application. 
We also build finite state transducers (FST) that take 
morphological tags as input and detect the NEs.
Manually built FSTs can express rules that are not intuitive (and sometimes impossible) 
to express with local grammars using regular expressions.
Consider, for example, counting the number of tags up to a threshold, or capturing left and right recursive structures 
which are common in Arabic since an Arabic phrase-sentence can start with a name or a verb.
We finally infer relations between the NEs based on their context.

@@

We evaluated our method and used it to solve two
NLP applications including extracting genealogical family
trees from Biblical text, and extracting a graph of 
narrators from hadith books.%, and extracting biographies from biography books.
Our results show high precision and recall metrics and report
learning interesting named entities and relational lemmas.
In this paper we make the following contributions.
\begin{itemize}
\item We present a method for named entity and entity graph
extraction from Arabic text using morphology, state 
machines, and graph transformation algorithms.
\item We present an Arabic lexicon augmented 
with (1) person names collected from several resources as well as
learned from our corpora, and (2) tags suitable to extract complex names,
kinship, and narration sequence relations. 
\item We learn semantics for morphological features associated 
with detected phrases that are equivalent to predefined features
modulo one or two graph transformations. 
The learned semantics improved recall and precision by 
5 and 6\%, respectively in the genealogy application.
\item We perform Arabic named entity and 
relational entity extraction using graph algorithms. 
%The use of a graph generated from the hadith application 
%improved precision in the biography application from 40\% to 93\%.
%\item We present a novel equality metric to compare Arabic 
%names based on name structure and morphological features. 
\end{itemize}

@@

Our method takes a set of Arabic text documents and builds 
a graph where named entities are nodes and edges are relations 
between the nodes. 
Our target graph has a predefined set of relations expressed 
as edge labels such as ``parent'' and ``spouse'' where the nodes
are person names.
We use Sarf, an inhouse morphological analyzer, augmented with 
a list of lexicon elements relevant to the application.
We also augment Sarf with 
a set of tags that classify the 
elements into several categories 
relevant to the targeted named entities and edge labels.

The method passes the input text string to Sarf.
Sarf processes the input text and whenever
it identifies a morpheme, it calls the application morpheme
classifier with the morpheme, the tags of the morpheme, 
and the current solution context.
Note that several morphological solutions may exist for the same
input string.
The morpheme classifier either stores the morpheme in a list of
unresolved morphemes, 
or resolves the unresolved morphemes and produces
a category as input to the application FST.
The application FST is manually built to
detect {\bf (1)} named entities,  
{\bf (2)} sequences of named entities,  
{\bf (3)} predefined edges connecting named entities, 
{\bf (4)} previously unknown named entities occurring between predefined edges, 
and {\bf (5)} previously unknown edge labels occurring between named entities.
%Note that there might be more than one level of finite state
%transducers to handle an NLP application.
The detected named entities and edges form an entity graph.

Finally, the method uses a distance
function to compute equality between the named entities and
to consequently merge the nodes of the graph that represent 
equal named entities.
%When nodes are merged, the edge labels are augmented to be the 
%concatenation of old edges.
The method then learns semantics for the discovered edge labels 
and suggests edge replacement algorithms to transform the 
discovered edge labels into the predefined edges. 
The method also interactively learns the semantics of other 
detected edge labels that it could not infer 
in terms of the predefined set of relations.

@@

%Due to religious and political reasons, 
%writing the traditions was forbidden 
%until the days of the Umayad Calif,
%\novocalize
%\RL{`mr bn `bd al`zyz}(717-720 AC), seventy or so years after the prophet. 
%Many inconsistencies were introduced to the literature which necessitates 
%a thorough authentication study before the use of a tradition in jurisprudence.
%\vocalize
%%An Islamic hadith scholar studies the authenticity of 
%%the \noArRL{sanad} of a 
%%set of related traditions before using these traditions 
%%jurisprudence. 
%%While different Islamic schools of thought differ on 
%%how to interpret the content, they almost all agree
%All Islamic schools of thought agree
%that scholars can not use a tradition whose sanad lacks authenticity for jurisprudence.
%
%The authenticity of a \noArRL{.hady_t} depends on 
%the credibility of the narrators as reported in 
%separate biography books. 
%The study of \noArRL{.hady_t} authentication is 
%currently manual and error prone due to the huge number
%of existing traditions and tradition books. 
%Hadithopedia~\cite{Hadithopaedia:08}
%reports that a non-comprehensive collection of tradition 
%and biography books amounts to more than 300 thousand lines of 
%text. 
Al-Azami\cite{Al-Azami-91} cites more than eleven books
of digitized tradition books each of several volumes, and a dozen
other biography and secondary authentication books such
as a geographical dictionary of places in hadith.

@@

The hadith FST generates sequences of hadith narrators
such as $\langle n_1,n_2,n_3,n_4,n_5 \rangle$ shown in 
Figure~\ref{f:exhadith}. 
Our target is to automate the extraction of 
narrator graphs,
similar to the directed acyclic graph (DAG) 
in Figure~\ref{f:hadith}(b),
from one or more narration books
%We automatically extracted the POR from the three books of 
%hadith~\cite{IbnHanbal,AlKulayni,AlTousi}.
The nodes in boxes are the \RL{matn} of the hadith, 
and the other nodes are the narrators.

@@

%\subsection{Narrator distance metric}
%MultiKey Hashing
%explain what are the keys used for each value and why suitable
% and say that we use the computed keys also conduct the 
% biography lookup
%Two narrator names may refer to the same person but may 
%differ in text and in structure. 
%For example, the names \noTrutfRL{عبد الله بن عمر}, 
%\noTrutfRL{ابن عمر}, and \noTrutfRL{ابن عمر رضي الله عنه}
%refer to the same person. 
%Our metric uses the \noTrutfRL{ابن} (son) name connector and 
%its morphological equivalents to order the person names
%that appear within a narrator entity. 
%The metric compares the name hierarchy and if the hierarchy 
%does not match, as in the case of 
%\noTrutfRL{عبيد الله بن عبد الله بن عمر}  and his parent 
%\noTrutfRL{عبد الله بن عمر},
%the metric returns \cci{false}.
%The metric also compares name qualifiers and if two name 
%qualifiers of the same morphological POS tag,  
%such as the place possessive qualifiers 
%\utfRL{العراقي} (the Iraqi) and \utfRL{المصري} (the Egyptian) 
%conflict, the metric returns \cci{false}.
%The metric then compares the morphological stems of the person 
%names that are in order, ignores the ones that are not in order,
%and computes the fraction of the matching names.
%
%Both our metric and the Levenshtein distance metric 
%used in~\cite{Azmi-2010}
%achieved 100\% precision on an arbitrarily selected set that 
%included 208 narrator pairs with 102 equal narrators.
%Our metric achieved 95\% recall versus 72\% recall for the 
%Levenshtein distance metric. 
%Which means that both metrics do not announce names equal if they
%are not, but our metric is better at detecting equal names.
%We believe that our metric is adequate to compare 
%complex Arabic names including contemporary ones in general.

%\section{The biography application}
%Biography books include one or more short biography 
%for each hadith narrator. 
%The short biography may contain information 
%about the dates and locations of birth and death, 
%the professors and the students of the narrator,
%and one or more evaluations of the credibility of the narrator. 
%
%We use a similar approach to the hadith application
%to locate narrator names within the biography books. 
%A biography text may contain the name of the narrator described
%in the biography as well as the names of other narrators who happen
%to be the professors, students, or the persons who evaluated the 
%narrator.
%Our method uses the narrator graph 
%extracted from several hadith books 
%as described earlier to segment the biographies
%and detect the owner of each biography. 
%
%Once our method detects a narrator, it colors the node in  the 
%narrator graph corresponding to it. 
%The professors and students of the narrator
%are the neighbors of the 
%node corresponding to the narrator in the graph. 
%Consequently, whenever a detected narrator entity 
%moves away from the 
%colored local cluster in the narrator graph, 
%we change the color and infer that we moved to a separate
%biography.
%The owner of each biography is the narrator at the center of the
%colored graph.
%
%The naive implementation of this approach is computationally
%expensive as it requires searching for each discovered narrator
%in the narrator graph. 
%We implemented a multi-hash narrator technique to 
%improve narrator localization in the graph and in the biography 
%books where we associate
%each narrator with all narrator names that can be generated 
%from its structure and compute a hash key and a relevance score 
%for that association. 
%We omit the discussion for brevity and we will report on that 
%elsewhere. 
%
%Scholars can use the located biographies to annotate
%each narrator with corresponding biographies thus automating
%the task of hadith authentication checks. 
%They can also annotate the narrator graph narrator graph with the locations and 
%the lifetime of the narrators and 
%check for time and location overlap that are necessary conditions
%for the consistency of two 
%neighboring narrators. 
%Scholars can also use the method to check the correspondence
%of different biography and hadith books. 
%
%We believe that this is the first work that applies
%{\em Arabic cross-document NLP}, and we believe
%that this approach can be extended to several applications
%such as the cross analysis of (1) investigation versus 
%interrogation reports, and (2) insurance claims versus 
%accident police reports to find inconsistencies.

@@

Our method takes as input Arabic text with a genealogical
list such as the genealogical list in
Figure~\ref{f:genelists}(a) 
and automatically extracts the genealogical family tree 
where person names are entities and edges represent 
``spouse'' and ``child'' labels such as 
Figure~\ref{f:genelists}(b).
We marked the person names with bars above them, and we 
filled the spouse nodes with grey. 
To perform this task, our method first passes the text to
Sarf, the morphological analyzer, that reports detected morphemes 
with their corresponding tags.

We augmented the tag set of Sarf with 
the following categories relevant to the genealogical 
application. 
\begin{itemize}
\item {\em NAME} a person name that can be either male or female. 509 lexicon items. 
\item {\em CHILD} a connector with a descent semantics such as \utfRL{ابن} (son). 12 lexicon items. 
\item {\em CHILDREN} a connector with a plural descent semantics such as \utfRL{ابناء} (children). 16 lexicon items. 
\item {\em PARENT} a connector with parenthood semantics such as \utfRL{أباه} (his father). 7 lexicon items. 
\item {\em SPOUSE} a connector with marital partnership semantics such as \utfRL{سريّته} (his concubine). 11 lexicon items. 
\item {\em SPOUSES} a connector with plural partnership semantics such as \utfRL{سراري} (his concubines). 7 lexicon items. 
\item {\em SIBLING} a connector with sisterhood semantics such as \utfRL{أخيه} (his brother). 14 lexicon items.
\item {\em SIBLINGS} a connector with plural sisterhood semantics such as \utfRL{اخوته} (his siblings). 18 lexicon items. 
\end{itemize}
Note that (1) some lexicon items might be categorized into 
more than one of the above categories, and 
(2) the method infers the gender from POS and gloss tags 
of the person name as well as from those of the 
other categories. 
In case of conflict, the method gives precedence to the 
gender inferred from the person names. 
Note also that the phrases
\utfRL{اسحاق بن يعقوب} (Izhac - son of - Jacob)
and 
\utfRL{اسحاق ابنه يعقوب} (Izhac - his son - Jacob)
contain the same named entities, and are identical 
as far as stemming is concerned. 
However, the analysis 
of the suffix morphemes of \utfRL{ابنه} results in inferring
a parent category in the second phrase.

@@

ANERSys boosts its capabilities to detect named entities by using 
task specific corpora and gazetteers and ignores 
POS tags~\cite{Benajiba:07,Benajiba:08}. 
We are similar to ANERSys in that we augment the lexicon Sarf
with application relevant lexicon items and tags. 
We differ in that they use statistical analysis models such as
n-gram and maximum entropy measures.
In other work, they
achieve major improvements when they use lexical and
morphosyntactic features and a parallel Arabic and English 
corpora to bootstrap noisy features~\cite{BenajibaZDR10}.

@@

We evaluated our method against three books of 
hadith~\cite{IbnHanbal,AlTousi,AlKulayni},
and the book of Genesis~\cite{}.
%and one book of biographies with several volumes~\cite{khoei}. 
We define our metrics, present our evaluation reference with interannotation agreement results, 
and then present the results for the hadith
and genealogical list applications.

@@

%
%{\bf Biography narrator detection.}~~
%Table~\ref{t:narratorDifference} shows the accuracy results for the narrator detection and the biography boundary detection
%for the Khoei biography book.
%The first row shows the results computed by considering the biography book alone without the help of the narrator graph
%extracted from the hadith books. 
%Without the narrator graph, our method detected 94\% of the narrators mentioned in the book with 81\% precision. 
%The method without the narrator graph, detected correctly (94\% precision) the boundaries of the biographies. 
%
%With the help of the narrator graph, our method detected 65\% of the narrator names in the biography book. 
%The rest of the names were simply not found in the hadith books that we used to generate the narrator graph. 
%With the use of the narrator graph and the coloring algorithm, our method detected the boundaries of the biographies
%with 93\% precision. 
%We believe that an exhaustive graph that includes narrator names from more books of hadith will improve the 65\% narrator 
%detection recall metric.
%In all cases, narrator names that exist in the biography and not in the hadith books are not of interest to the narrator
%graph annotation problem.
%
%%The numbers in the right hand side of Table~\ref{t:narratorDifference}  may explain why 
%Narrator detection in biography books is harder than in hadith books. 
%For example, the ratio of narrator names to the rest of the text is 22\% in Musnad while it is 49\% in Khoei,
%and the average length of a narrator sequence in Musnad is 2.6 narrators while it is 3.5 narrators in Khoei.
%With the use of the narrator graph extracted from the hadith books, our method was able to improve biography boundary detection
%from 41\% to 93\%.
%This is evidence that cross-document entity detection helps to accurately find named entities in harder NLP tasks. 
%
%\begin{table*}[bt]                                     
%\centering                                            
%\caption{Accuracy results for the biography application.}
%\resizebox{.7\columnwidth}{!}{
%
%%\begin{tabular}{cp{0.3cm}c}
%
%	\begin{tabular}{cp{.1cm}cccp{.1cm}ccc}
%	          & & \multicolumn{3}{c}{narrator detection} & &  \multicolumn{3}{c}{biography boundary} \\
%	 & & recall & precision & F-score & & recall & precision & F-score \\ 
% \cline{3-5} \cline{7-9}
%%hadith	 & sanad & & 0.923 & 0.999 & 0.960 & & 0.996 & 0.999 & 0.998  \\
%%(musnad) & all & & 0.954 & 0.961 & 0.958 & & 0.996 & 0.987 & 0.992 \\ 
%%	\hline \hline
%w/o graph & & 0.943 & 0.812 & 0.873 & & 0.407 & 0.945 & 0.569 \\
%with graph & & 0.654 & 0.909 & 0.760 & & 0.933 & 0.960 & 0.946 \\ % found in narrator graph
%	 \end{tabular}   
%%&  &
%%	\begin{tabular}{ccc}
%%	& names    & nar. \\
%%    & per text & elements \\ \cline{2-3}
%%	\multirow{2}{*}{0.217} &\multirow{2}{*}{ 2.64}\\ \\
%%	\hline \hline
%%	\multirow{2}{*}{0.489} & \multirow{2}{*}{3.54}\\
%%    musnad & 0.217 & 2.64 \\ 
%%    biography & 0.489 & 3.54 \\
%%	\end{tabular}  \\
%%\end{tabular}          
%}                             
%\normalsize                   
%\label{t:narratorDifference} 
%\end{table*}    
%
